```{r, include=FALSE}
# The 'include=FALSE' hides the code in the output file.

# load data
iods <- read.table("data/iods.txt", header = T)
```

# Chapter title

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

## Including R output

Include R code and output such as graphs and tables inside 'code chunks' like below:  

```{r, warning=FALSE}
library(ggplot2)
qplot(Attitude, Points, data = iods) + geom_smooth(method = "lm")
```


## Helpful resources

- RMarkdown [cheat sheat](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
- RMarkdown [reference guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)  
- RMarkdown HTML documents [documentation](http://rmarkdown.rstudio.com/html_document_format.html)
- In the index.Rmd file you can choose a [bootswatch theme](http://bootswatch.com/) for you page