```{r, include=FALSE}
# load data
iods <- read.table("data/iods.txt", header = T)
```

# Project title

*Describe the project here and summarize your learning.*

- Describe your work and results clearly  
- Assume the reader has an introductory course level understanding of writing and reading R code  
- Assume the reader has an introductory course level understanding of statistical methods  
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

### Including R output

Include R code and output such as graphs and tables inside 'code chunks' like below:  

```{r}
library(ggplot2)
qplot(Attitude, Points, data = iods) + geom_smooth(method = "lm")
```


### Helpful resources

- Rmarkdown [cheat sheat](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
- HTML documents [Documentation](http://rmarkdown.rstudio.com/html_document_format.html)
- in the index.Rmd you can choose on of [these themes](http://bootswatch.com/) for you page